\plaintitle

### 2016-05-27 cscw/papers/2017/labor/trackingWork.tex
refining intro

``But'',
as \citeauthor{scholz2012digital} points out,
``it would be wrong to conclude that
in the realm of digital labor there is nothing new under the sun''
\cite{scholz2012digital}.
For one, information work
(e.g. AMT \& Upwork)
and digitally managed work in general make it substantially easier
to keep workers geographically dispersed
compared to the factory workers that followed home--based piecework.
This aspect of work renders workers virtually invisible to the task solicitors
--- or ``requesters'' ---
despite being significantly more connected than historical pieceworkers were % this is a weak point...
\cite{turkopticon}.
Secondly, while
factory foremen and other middle--men
traditionally mediated the relationships between workers and managers,
today the visible agents are \textit{systems};
rather than employing individuals who ``personified the functions of management''
and can thus negotiate workers' needs,
socio--technical systems mediate these interactions % often with unrewarding outcomes
\cite{wray1949marginal}.
% \cite{storiesIraniSilberman}.
Finally, this characteristic of modern piecework makes
worker coordination for collective advocacy and action significantly more difficult,
necessitating special consideration to maintain the inertia of collective will
while focusing that energy productively
\cite{dynamo}.
\ali{I really don't like this paragraph\dots}
% here are undoubtedly more examples of differences.
% How do I not stake a claim that this is comprehensive?
% Or should I really flesh this out and make it (comprehensive)?

\subsection{Piecework as a lens to understand gig work}
\citeauthor{crowdworkFuture}
investigate the future of crowd work by
situating and interrogating it through ``piecework'',
which historically described work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture}.
% \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
% This work was largely in textiles and was predominantly done by women,
% which by itself is a topic worthy of unpacking
% \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
% \ali{let's just leave this in the graveyard for now; i feel like someone who knows what they're talking about with regard to feminist and gender studies
% (e.g. the Bardzells, Irani) would really do justice here, but i'd be getting in over my head}
% \cite{scott1975women}.
Given the scope,
we can frame piecework and gig work as sharing these important similarities:
\begin{inlinelist}
\item this form of work began in the home
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
% \msb{my first reaction was ``this isn't definitional''. But now I look back and see that you're just saying there are similarities. It would be helpful if the sentence made it clearer that the similarities were in not just what the work was but also in the environment around it.}
\end{inlinelist}.

\subsection{Context}
Importantly, since the reemergence of piecework we've seen substantial frustration
and resistance among the workers in this area
\cite{uberAlgorithm,turkopticon,dynamo}.
This paper attempts to make sense of the broader research on this piecework,
or ``gig work'', by framing this as one of several steps in the marginalization of workers,
% \msb{this feels out of order to me. I don't think resistance is a step in the marginalization of workers, isn't it a reaction to the marginalization of researchers?}
%
% this needs to be re-written to reflect the new split
%
%
% starting with the discretization of tasks,
% followed by routinization and the rise of work--flows,
% and finally the external management of workers.
% 
All of this is to say that these milestones follow sequentially,
not coincidentally but necessarily,
and that by tracing this path using the corpus of scholarship on labor and workers
we can both make sense of past events and perhaps reasonably predict next steps.
% \msb{I don't think we've offered much proof that this is the case yet. Why must they follow necessarily? Can you offer a cause and effect relationship?}

We explore each topic
--- discretization, routinization, \& management ---
by looking at case studies in social computing.
Having validated this lens as a way of reasoning about contemporary piecework, or ``gig work'',
we turn to look ahead, envisioning future areas that researchers in social computing
--- and particularly digitally mediated work ---
should explore.

\section{Case Studies}
The existing body of research has shed light on on--demand from various perspectives,
and revealed a number of topics that,
through our framing, are clearly situated together.
Those topics are, at a high level, as follows:
\begin{enumerate}
\item the processes involved in making work into tasks, or \textbf{discretization};
% in ``\nameref{sec:MakingGigsWork}'';
\item the outcomes of that discretization, both on the work itself as well as the workers;
% in ``\nameref{sec:Fallout}'';
and finally
\item the results of these forces; namely the turning of work toward \textbf{factories}.
% which we will discuss in ``\nameref{sec:Factorization}''.
\end{enumerate}

We will follow these topics, using them as prompts for case studies in
the emergence and
development of contemporary piecework.

@@

\subsubsection{Decomposition}\label{sec:decomposition}
Piecework may be thought of as vertically slicing work such that 
each person is responsible for the whole task
--- making a whole garment, in this case.
Broken down in this way, work could grow to unprecedented scales,
but the quality of the work would remain relatively variable.
Textile work being a salient example,
it took time for workers to acquire sufficient skill
to do every aspect of the piece work so that the garment would be accepted.
% some seamstresses might be better than others at the same task,
% making this framing of the work initially problematic.

>>>>

``But'',
as \citeauthor{scholz2012digital} points out,
``it would be wrong to conclude that
in the realm of digital labor there is nothing new under the sun''
\cite{scholz2012digital}.
For one, information work
(e.g. AMT \& Upwork)
and digitally managed work in general make it substantially easier
to keep workers geographically dispersed
compared to the factory workers that followed home--based piecework.
This aspect of work renders workers virtually invisible to the task solicitors
--- or ``requesters'' ---
despite being significantly more connected than historical pieceworkers were % this is a weak point...
\cite{turkopticon}.
Secondly, while
factory foremen and other middle--men
traditionally mediated the relationships between workers and managers,
today the visible agents are \textit{systems};
rather than employing individuals who ``personified the functions of management''
and can thus negotiate workers' needs,
socio--technical systems mediate these interactions % often with unrewarding outcomes
\cite{wray1949marginal}.
% \cite{storiesIraniSilberman}.
Finally, this characteristic of modern piecework makes
worker coordination for collective advocacy and action significantly more difficult,
necessitating special consideration to maintain the inertia of collective will
while focusing that energy productively
\cite{dynamo}.
% \ali{I really don't like this paragraph\dots}
% here are undoubtedly more examples of differences.
% How do I not stake a claim that this is comprehensive?
% Or should I really flesh this out and make it (comprehensive)?

\subsection{Piecework as a lens to understand gig work}

\citeauthor{crowdworkFuture}
write of the future of crowd work in an effort to investigate
whether crowd work will be a worthwhile form of work or another instantiation of ``piecework'',
a term historically used to describe work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture}.
% \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
% This work was largely in textiles and was predominantly done by women,
% which by itself is a topic worthy of unpacking
% \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
% \ali{let's just leave this in the graveyard for now; i feel like someone who knows what they're talking about with regard to feminist and gender studies
% (e.g. the Bardzells, Irani) would really do justice here, but i'd be getting in over my head}
% \cite{scott1975women}.
Given the scope,
we can frame piecework and gig work as sharing these important similarities:
\begin{inlinelist}
\item this form of work began in the home
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
\end{inlinelist}.

% \subsection{Context}
In the past decade, researchers have observed frustration
grow among on--demand workers,
with expression of this frustration spanning a wide range of tactics
\cite{uberAlgorithm,turkopticon,dynamo}.
Attempting to make sense of these case studies has been challenging
in part because
a wholly encompassing framework for understanding this form of work
has thus proven difficult to capture.

This paper attempts to make sense of the broader research on this new form of labor,
or ``gig work'', by evaluating this work through a more familiar lens: piecework.
More concretely, by looking at task--based or ``gig'' work as
an instantiation and continuation of piecework,
and by looking for patterns of behavior that the corresponding literature predicts
on this basis, we can do the following:
\begin{inlinelist}
  \item we can make sense of the phenomena so far as part of a much larger series of interrelated events;
  \item we can bring into focus the ongoing work among workers, system--designers, and researchers in this space; and finally,
  \item we can offer predictions of what social computing researchers, and workers themselves, should expect to see on the horizon of on--demand work.
\end{inlinelist}

We'll look at a broad range of cases under a number of major themes
we propose as broadly describing the types of research being done in crowd work
and more generally in what we argue is contemporary piecework.
After validating this lens as a way of reasoning about on--demand labor,
we'll attempt to use this perspective to suggest areas of research worth anticipating,
and developments we should expect to see in the maturation of digitally mediated work.

\section{Case Studies}
The existing body of research has shed light on on--demand from various perspectives,
and revealed a number of topics that,
through our framing, are clearly situated together.
Those topics are, at a high level, as follows:
\begin{enumerate}
\item the processes involved in making work into tasks, or \textbf{discretization};
\item the outcomes of that discretization, both on the work itself as well as the workers;
and finally
\item the results of these forces; namely the turning of work toward \textbf{factories}.
\end{enumerate}

% We will follow these topics, using them as prompts for case studies in
% the emergence and
% development of contemporary piecework.

@@

\subsubsection{Decomposition}\label{sec:decomposition}
Piecework may be thought of as vertically slicing work such that 
each person is responsible for the whole task
--- making a whole garment, in this case.
Broken down in this way, work could grow to unprecedented scales,
but the quality of the work would remain relatively variable.
Textile work being a salient example,
it took time for workers to acquire sufficient skill
to do every aspect of the work so that the garment would be accepted by the company soliciting that work.
% some seamstresses might be better than others at the same task,
% making this framing of the work initially problematic.
